ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-07-24 pyked.tex
edits through beginning of section 2

Instead, such data is typically found in under-documented comma-separated value
(CSV) files and Excel spreadsheets, or contained in PDF tables rather than as
supplementary material associated with a paper. In the worst case, data is only
available in figures and must be digitized by the use of software such as
WebPlotDigitizer~\autocite{WebPlotDigitizer}. These practices limit the wide use
of the valuable, extensive experimental data available in the literature.
\textcite{Frenklach:2007bm} further explained the benefits of standardized,
widely available data for combustion. In brief, much of fundamental
combustion\slash chemical kinetics research drives towards the ultimate goal of
developing predictive kinetic models---and ultimately this depends on a
\emph{community} infrastructure for data and methods.

@@

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility specific
effects important to properly simulating certain experiments. Second, the XML
format is intended to be a markup language rather than a data format, and
presents a barrier to creating and working with database files. Furthermore, the
focus on using internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products and InCHI or SMILES for species), limits usability outside
the PrIMe ecosystem. Finally, the closed and opaque nature of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the
ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
features that make files better standalone representations of experimental
data---i.e., more informative by themselves rather than in concert with a larger
system. For example, where PrIMe uses internal bibliographic references,
ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
provides machine-readable formats for specifying ignition experiments, including
a field for the definition of autoignition and the ability to specify
facility-specific effects. However, ReSpecTh experimental files are another
XML-based format, and, as such, suffer from the same usability issues as PrIMe.
Moreover, the standard remains closed, requiring potential users to register
with the standard's authors to access the specification.

@@

Despite the development of these standards, limited examples of open data
\emph{sharing} using the formats exist in the literature. To the best
of our knowledge, the largest such publicly available database is hosted by
the Clean Combustion Research Center at King Abdullah University of Science and
Technology (KAUST) and is known as CloudFlame
(\url{https://cloudflame.kaust.edu.sa}). Started in
2013~\autocite{Goteng:2013cf,Goteng:2014,ReynoChiasson:2015}, CloudFlame serves
as an openly accessible database for experimental data, available in a standard
CSV format, and also provides a cloud infrastructure for running simulations
based on stored models and data. While an admirable and useful effort,
CloudFlame, like ReSpecTh and PrIMe, is a closed system controlled by a single
institution rather than by the community at large.

@@

This section will provide a brief introduction to the current version of the
\ck{} database format, v0.1.6, which represents a functional standard
for documenting autoignition measurements in shock tubes and rapid compression machines.
This is not a complete listing of available fields and a complete reference for the
schema can be found in the online documentation.
\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
In addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
files. We indicate \yabox|yaml| keywords or values using text with gray background,
and \pybox|Python| code using text with a yellow background.

@@

\ck{} files are stored online in a git repository hosted on GitHub.
\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
Submissions from the community are welcomed and can be made via the standard Pull Request
on GitHub. The authors also encourage questions and problems to be submitted either via
email or as Issues on GitHub. In particular, researchers interested in converters for
internal data formats to the \ck{} format are encouraged to reach out. In addition,
the \pk{} package (described in more detail in Sec.~\label{sec:pyked-architecture})
provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.

>>>>

Instead, such data is typically found in under-documented comma-separated value
(CSV) files and Excel spreadsheets, or contained in PDF tables rather than as
supplementary material associated with a paper. In the worst case, data is only
available in figures and must be digitized by the use of software such as
WebPlotDigitizer~\autocite{WebPlotDigitizer}. These practices limit wide use
of the valuable, extensive experimental data available in the literature.
\textcite{Frenklach:2007bm} further explained the benefits of standardized,
widely available data for combustion. In brief, much of fundamental
combustion\slash chemical kinetics research drives towards the ultimate goal of
developing predictive kinetic models---and ultimately this depends on a
\emph{community} infrastructure for data and methods.

@@

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a markup language rather than a data format, and
its lack of human readibility presents a barrier to creating and working with
database files. Furthermore, the
focus on using internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species), limits usability outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the
ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
features that make files better standalone representations of experimental
data---i.e., more informative by themselves rather than in concert with a larger
system. For example, where PrIMe uses internal bibliographic references,
ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
provides machine-readable formats for describing ignition experiments, including
a field for the definition of autoignition and the ability to specify
facility-specific effects. However, ReSpecTh experimental files are another
XML-based format, and, as such, suffer from the same usability issues as PrIMe.
Moreover, the standard remains closed, requiring potential users to register
with the standard's authors to access the specification and the growing database.

@@

Despite the development of these standards, limited examples exist in the
literature of open data \emph{sharing} using the formats. To the best
of our knowledge, the largest such publicly available database is hosted by
the Clean Combustion Research Center at King Abdullah University of Science and
Technology (KAUST): CloudFlame.\footnote{\url{https://cloudflame.kaust.edu.sa}}
Started in 2013~\autocite{Goteng:2013cf,Goteng:2014,ReynoChiasson:2015}, CloudFlame
serves as an openly accessible database for experimental data, available in a standard
CSV format, and also provides a cloud infrastructure for running simulations
based on stored models and data. While an admirable and useful effort,
CloudFlame, like ReSpecTh and PrIMe, is a closed system controlled by a single
institution rather than by the community at large.

@@

This section will provide a brief introduction to the current version of the
\ck{} database format, v0.1.6, which represents a functional standard
for documenting autoignition measurements in shock tubes and rapid compression machines.
Here we do not include a complete listing of available fields; a complete reference for the
schema can be found in the online documentation.
\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
In addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
files. We indicate \yabox|yaml| keywords or values using text with gray background,
and \pybox|Python| code using text with a yellow background.

@@

\ck{} files are stored online in a git repository hosted on GitHub.
\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
Submissions from the community are welcomed and can be made via the standard Pull Request
on GitHub. The authors also encourage questions and problems to be submitted either via
email or as Issues on GitHub. In particular, researchers interested in converters for
internal data formats to the \ck{} format are encouraged to reach out. In addition,
the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
